Unicode Implementation Issues \thanks { This is a working document for the Rexx Language Association (RexxLA). There will be one or more separate requirements documents. It is written in \LaTeX\ \fmtversion\ and was rendered \today\ using \enginedetails. }

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/evaluation.tex
add filenames

This section refers to the DATATYPE built-in function when checking
operands, see nnn. Except for considerations of limits on the values of
exponents, the test:

datatype (Subject) == `NUM' is equivalent to testing whether the subject
matches the syntax: num := {[}blank+{]} {[}`+' \textbar{} `-'{]}
{[}blank+{]} number {[}blank+{]}

For the syntax of number see nnn.

When the matching subject does not include a `-' the value is the value
of the number in the match, otherwise the value is the value of the
expression (0 - number).

@@

datatype (Subject , `W')

is a test that the Subject matches that syntax and also has a value that
is ``whole'', that is has no non-zero fractional part.

When these two tests are made and the Subject matches the constraints
but has an exponent that is not

in the correct range of values then a condition is raised: call \#Raise
`SYNTAX', 41.7, Subject

This possibility is implied by the uses of DATATYPE and not shown
explicitly in the rest of this section nnn.

@@

The notation used here is the same as that defined in sections nnn and
nnn, including the fact that the Var\_ routines may return an indicator
of `N', `S' or `X'.

Each possible name in a variable pool is qualified as tailed or
non-tailed name; names with different qualification and the same
spelling are different items in the pool. For those Var\_ functions with
a third argument this argument indicates the qualification; it is `1'
when addressing tailed names or `0' when addressing non-tailed names.

@@

A variable pool is associated with a reference denoted by the first
argument, with name Pool. The value of Pool may alter during execution.
The same name, in conjunction with different values of Pool, can
correspond to different values.

@@

Var\_Empty (Pool)

@@

Var \_Set(Pool, Name, `0', Value)

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then Var\_Set operates on the variable pool referenced by
\#Upper in this pool and this rule is applied to that pool. When the
pool with attribute `not-exposed' for this name is determined the
specified value is associated with the specified name. It also
associates the attributes `not-dropped' and `not-implicit'. If that
attribute was previously `not-dropped' then the indicator returned is
`R'. The name is a stem if it contains just one period, as its rightmost
character. When the name is a stem Var\_Set(Pool, TailedName, `1',Value)
is executed for all possible valid tailed names which

have Name as their stem, and then those tailed-names are given the
attribute `implicit'. Var \_Set(Pool, Name, `1', Value)

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then Var\_Set operates
on the variable pool referenced by \#Upper in this pool and this rule is
applied to that pool. When the pool with attribute `not-exposed' for the
stem is determined the name is considered in that pool. If the name has
the `exposed' attribute then the

variable pool referenced by \#Upper in the pool is considered and this
rule applied to that pool. When the pool with attribute `not-exposed' is
determined the specified value is associated with the specified name. It
also associates the attributes `not-dropped' and `not-implicit' . If
that attribute was previously `not-dropped' then the indicator returned
is `R'.

@@

Var \_Value(Pool, Name, `0')

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then Var\_Value operates on the variable pool referenced by
\#Upper in this pool and this rule is applied to that pool. When the
pool with attribute `not-exposed' for this name is determined the
indicator returned is `D' if the name has `dropped' associated, `N'
otherwise. In the former case \#Outcome is set equal to Name, in the
latter case \#Outcome is set to the value most

recently associated with the name by Var\_Set. Var \_Value(Pool, Name,
`1')

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then Var\_Value
operates on the variable pool referenced by \#Upper in this pool and
this rule is applied to that pool. When the pool with attribute
`not-exposed' for the stem is determined the name is considered in that
pool. If the name has the `exposed' attribute then the variable pool
referenced by \#Upper in the pool is considered and this rule applied to
that pool. When the pool with attribute `not-exposed' is determined the
indicator returned is `D' if the name has `dropped' associated, `N'
otherwise. In the former case \#Outcome is set equal to Name, in the
latter case \#Outcome is set to the value most recently associated with
the name by Var\_Set.

@@

Var \_Drop(Pool, Name, `0')

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then Var\_Drop operates on the variable pool referenced by
\#Upper in this pool and this rule is applied to that pool. When the
pool with attribute `not-exposed' for this name is determined the
attribute `dropped' is associated with the specified name. Also, when
the name is a stem, Var\_Drop(Pool,TailedName,`1') is executed for all
possible valid tailed names which have Name as astem.

Var \_Drop(Pool, Name, `1')

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then Var\_Drop
operates on the variable pool referenced by \#Upper in this pool and
this rule is applied to that pool. When the pool with attribute
`not-exposed' for the stem is determined the name is considered in that
pool. If the name has the `exposed' attribute then the variable pool
referenced by \#Upper in the pool is considered and this rule applied to
that pool. When the pool with attribute `not-exposed' is determined the
attribute `dropped' is associated with the specified name.

@@

Var\_Expose (Pool, Name, `0')

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. The attribute `exposed' is associated
with the specified name. Also, when the name is a stem,
Var\_Expose(Pool, TailedName,`1') is executed for all possible valid
tailed names which have Name as a stem.

Var\_Expose (Pool, Name, `1')

@@

Var\_ Reset (Pool)

The function operates on the variable pool with the specified reference.
It establishes the effect of subsequent API\_Next and API\_NextVariable
functions (see sections nnn and nnn). A Var\_Reset is implied by any
API\_ operation other than API\_Next and API\_NextVariable.

@@

The value of a symbol which is a NUMBER or a CONST\_SYMBOL which is not
a reserved symbol is the content of the appropriate token.

The value of a VAR\_SYMBOL which is ``taken as a constant'' is the
VAR\_SYMBOL itself, otherwise the VAR\_SYMBOL identifies a variable and
its value may vary during execution.

Accessing the value of a symbol which is not ``taken as a constant''
shall result in trace output, see nnn: if \#Tracing.\#Level == `I' then
call \#Trace Tag

where Tag is `\textgreater L\textgreater{}' unless the symbol is a
VAR\_SYMBOL which, when used as an argument to Var\_Value, does not
yield an indicator `D'. In that case, the Tag is
'\textgreater V\textgreater``.

@@

If VAR\_SYMBOL does not contain a period, or contains only one period as
its last character, the value of

the variable is the value associated with VAR\_SYMBOL in the variable
pool, that is \#Outcome after Var\_ Value (Pool, VAR SYMBOL, `0')

If the indicator is `D', indicating the variable has the `dropped'
attribute, the NOVALUE condition is raised; see nnn and nnn for
exceptions to this. \#Response = Var Value(Pool, VAR SYMBOL, `0') if
left(\#Response,1) == `D' then call \#Raise `NOVALUE', VAR\_SYMBOL, '\,'
If VAR\_SYMBOL contains a period which is not its last character, the
value of the variable is the value associated with the derived name.
7.3.1 Derived names A derived name is derived from a VAR\_SYMBOL as
follows:

VAR SYMBOL := Stem Tail

Stem := PlainSymbol `.'!

Tail i= (PlainSymbol \textbar{} `.' {[}PlainSymbol{]}) {[}`.'
{[}PlainSymbol1{]}{]}+ PlainSymbol := (general letter \textbar{} digit)+

The derived name is the concatenation of: - the Stem, without further
evaluation; - the Tail, with the PlainSymbols replaced by the values of
the symbols. The value of a PlainSymbol which does not start with a
digit is \#Outcome after Var\_ Value (Pool, PlainSymbol,`0') These
values are obtained without raising the NOVALUE condition.

@@

If the comparison has a comparison\_operator that is a normal\_compare
then the variable \#Test is set as follows:
\lstinputlisting[language=rexx,label=evaluation-comparison.rexx,caption=evaluation-comparison.rexx]{evaluation-comparison.rexx}

The value of \#Test, in conjunction with the operator in the comparison,
determines the value of the comparison. The value of the comparison is
`1' if - \#Test is `E' and the operator is one of `=``\,``,'==````,
`\textgreater=', \textless='', `\textgreater{}', `\textless{}',
`p\textgreater=', `\textless\textless=', \textgreater\textgreater', or
\textless\textless)

\begin{itemize}
\tightlist
\item
  \#Test is `G' and the operator is one of `\textgreater{}',
  `\textgreater=``, `\textless{}', `=', `\textless\textgreater{}',
  `\textgreater\textless{}', Nes'', `\textgreater\textgreater!
  `p\textgreater{}', or \textless\textless``)
\item
  \#Test is `L' and the operator is one of `\textless{}', \textless=``,
  \textgreater{}`, =', `\textless\textgreater{}',
  `\textgreater\textless{}', ==`,'\textless\textless',
  *\textless\textless=`, or \textgreater\textgreater{}'. In all other
  cases the value of the comparison is `0'.
\end{itemize}

@@

The code of ArithOp itself is assumed to operate under a sufficiently
high setting of numeric digits to avoid exponential notation.
\lstinputlisting[language=rexx,label=evaluation-arithmetic.rexx,caption=evaluation-arithmetic.rexx]{evaluation-arithmetic.rexx}

@@

Invocation occurs when a function or a message\_term or a callis
evaluated. Invocation of a function may result in a value, in which
case:

if \#Tracing.\#Level == `I' then call \#Trace
`\textgreater F\textgreater{}'

Invocation of a message\_term may result in a value, in which case:

if \#Tracing.\#Level == `I' then call \#Trace
`\textgreater M\textgreater{}'

@@

The argument positions are the positions in the exoression\_list where
syntactically an expression occurs or could have occurred. Let ArgNumber
be the number of an argument position, counting from 1 at the left; the
range of ArgNumber is all whole numbers greater than zero.

For each value of ArgNumber, \#ArgExists.\#NewLevel.ArgNumber is set `1'
if there is an expression present, `O' if not.

From the left, if \#ArgExists.\#NewLevel.ArgNumber is `1' then
\#Arg.\#NewLevel.ArgNumber is set to the value of the corresponding
expression. If \#ArgExists.\#NewLevel.ArgNumber is `0' then
\#Arg.\#NewLevel.ArgNumber is set to the null string.

\#ArgExists.\#NewLevel.0 is set to the largest ArgNumber for which
\#ArgExists.\#NewLevel.ArgNumber is `1', or to zero if there is no such
value of ArgNumber.

@@

The value of a LABEL, or of the taken\_constant in the function or
call\_instruction, is taken as a constant, see nnn. If the
taken\_constant is not a string\_literal it is a reference to the first
LABEL in the program which has the same value. The comparison is made
with the `==' operator.

If there is such a matching label and the label is trace-only (see nnn)
then a condition is raised: call \#Raise `SYNTAX', 16.3, taken constant

@@

If there is no such matching label, or if the taken\_constant is a
string\_literal, further comparisons are made.

If the value of the taken\_constant matches the name of some built-in
function then that built-in function is invoked. The names of the
built-in functions are defined in section nnn and are in uppercase.

If the value does not match any built-in function name,
Config\_ExternalRoutine is used to invoke an external routine.

Whenever a matching label is found, the variables SIGL and .SIGL are
assigned the value of the line number of the clause which caused the
search for the label. In the case of an invocation resulting from a

condition occurring that shall be the clause in which the condition
occurred. Var \_ Set(\#Pool, `SIGL', `0', \#LineNumber) var Set(0 ,
`.SIGL', `0', \#LineNumber)

The name used in the invocation is held in \#Name.\#Level for possible
use in an error message from the RETURN clause, see nnn

@@

An internal routine completes when \#Level returns to the value it had
when the routine was invoked. The value of the internal function is the
value of the expression on the return which completed the routine. The
value of an external function is determined by Config\_ExternalRoutine.

@@

An internal method completes when \#Level returns to the value it had
when the routine was invoked. The value of the internal method is the
value of the expression on the return which completed the method. The
value of an external method is determined by Config\_ExternalMethod.

@@

See nnn for the syntax of a message\_term. The value of the ferm within
a message\_term is called the receiver.

The receiver and any arguments of the term are evaluated, in left to
right order. r= \#evaluate(message term, term) If the message term
contains `\textasciitilde\textasciitilde{}' the value of the message
term is the receiver. Any effect on .Result? Otherwise the value of a
message\_term is the value of the method it invokes. The method invoked
is determined by the receiver and the taken\_constant and symbol. t =
\#Instance(message term, taken constant) If there is a symbol, it is
subject to a constraints. if \#contains (message term, symbol) then do
if r \textless\textgreater{} \#Self then

call \#Raise `SYNTAX', nn.n

/* OOI: ``Message search overrides can only be used from methods of the
target object.'' */

The search will progress from the object to its class and superclasses.
/* This is going to be circular because it describes the message lookup
algorithm and also uses messages. However for the messages in this code
the message names are chosen to be unique to a method so there is no
need to use this algorithm in deciding which method is intended. */

/* message term ::= receiver `\textasciitilde{}' taken constant `:'
VAR\_SYMBOL arguments */

/* This code reflects OOI - the arguments on the message don't affect
the method choice. */

/* This code selects a method based on its arguments, receiver,
taken\_constant, and symbol. */

/* This code is used in a context where \#Self is the receiver of the
method invocation which the subject message term is running under. */
\lstinputlisting[language=rexx,label=evaluation-selectmethod,caption=evaluation-selectmethod]{evaluation-selectmethod}

@@

The values of the arguments to the use of Config\_ExternalRoutine, in
order, are:

The argument How is `SUBROUTINE' if the invocation is from a call,
```FUNCTION' if the invocation is from a function.

The argument NameType is `1' if the taken\_constant is a
string\_literal, `0' otherwise.

The argument Name is the value of the faken\_constant.

The argument Environment is the value of this argument on the API\_
Start which started this execution. The argument Arguments is the \#Arg.
and \#ArgExists. data.

The argument Streams is the value of this argument on the API\_Start
which started this execution.

The argument Traps is the value of this argument on the API\_Start which
started this execution. Var\_Reset is invoked and \#API\_Enabled set to
`1' before use of Config\_ExternalRoutine. \#API\_Enabled is set to `O'
after.

The response from Config\_ExternalRoutine is processed. If no conditions
are (implicitly) raised, \#Outcome is the value of the function.

>>>>

This section refers to the \texttt{DATATYPE} built-in function when
checking operands, see nnn. Except for considerations of limits on the
values of exponents, the test:

\lstinputlisting[language=rexx,label=datatypenum.rexx,caption=datatypenum.rexx]{datatypenum.rexx}

is equivalent to testing whether the subject matches the syntax:

\lstinputlisting[language=rexx,label=ebfdatatype.ebnf,caption=ebfdatatype.ebnf]{ebfdatatype.ebnf}

For the syntax of \emph{number} see nnn.

When the matching subject does not include a
\texttt{\textquotesingle{}-\textquotesingle{}} the value is the value of
the number in the match, otherwise the value is the value of the
expression \texttt{(0\ -\ number)}.

@@

\lstinputlisting[language=rexx,label=datatypesubject.rexx,caption=datatypesubject.rexx]{datatypesubject.rexx}

is a test that the \texttt{Subject} matches that syntax and also has a
value that is ``whole'', that is has no non-zero fractional part.

When these two tests are made and the \texttt{Subject} matches the
constraints but has an exponent that is not in the correct range of
values then a condition is raised:

\lstinputlisting[language=rexx,label=callraisesyntax.rexx,caption=callraisesyntax.rexx]{callraisesyntax.rexx}

This possibility is implied by the uses of \texttt{DATATYPE} and not
shown explicitly in the rest of this section nnn.

@@

The notation used here is the same as that defined in sections nnn and
nnn, including the fact that the \texttt{Var\_} routines may return an
indicator of \texttt{\textquotesingle{}N\textquotesingle{}},
\texttt{\textquotesingle{}S\textquotesingle{}} or
\texttt{\textquotesingle{}X\textquotesingle{}}.

Each possible name in a variable pool is qualified as tailed or
non-tailed name; names with different qualification and the same
spelling are different items in the pool. For those \texttt{Var\_}
functions with a third argument this argument indicates the
qualification; it is \texttt{\textquotesingle{}1\textquotesingle{}} when
addressing tailed names or
\texttt{\textquotesingle{}0\textquotesingle{}} when addressing
non-tailed names.

@@

A variable pool is associated with a reference denoted by the first
argument, with name \texttt{Pool}. The value of \texttt{Pool} may alter
during execution. The same name, in conjunction with different values of
\texttt{Pool}, can correspond to different values.

@@

\lstinputlisting[language=rexx,label=varpoolempty.rexx,caption=varpoolempty.rexx]{varpoolempty.rexx}

@@

\lstinputlisting[language=rexx,label=varsetzero.rexx,caption=varsetzero.rexx]{varsetzero.rexx}

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then \texttt{Var\_Set} operates on the variable pool
referenced by \texttt{\#Upper} in this pool and this rule is applied to
that pool. When the pool with attribute `not-exposed' for this name is
determined the specified value is associated with the specified name. It
also associates the attributes `not-dropped' and `not-implicit'. If that
attribute was previously `not-dropped' then the indicator returned is
\texttt{\textquotesingle{}R\textquotesingle{}}. The name is a stem if it
contains just one period, as its rightmost character. When the name is a
stem
\texttt{Var\_Set(Pool,\ TailedName,\ \textquotesingle{}1\textquotesingle{},Value)}
is executed for all possible valid tailed names which have Name as their
stem, and then those tailed-names are given the attribute `implicit'.

\lstinputlisting[language=rexx,label=varsetone.rexx,caption=varsetone.rexx]{varsetone.rexx}

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then Var\_Set operates
on the variable pool referenced by \#Upper in this pool and this rule is
applied to that pool. When the pool with attribute `not-exposed' for the
stem is determined the name is considered in that pool. If the name has
the `exposed' attribute then the variable pool referenced by \#Upper in
the pool is considered and this rule applied to that pool. When the pool
with attribute `not-exposed' is determined the specified value is
associated with the specified name. It also associates the attributes
`not-dropped' and `not-implicit' . If that attribute was previously
`not-dropped' then the indicator returned is
\texttt{\textquotesingle{}R\textquotesingle{}}.

@@

\lstinputlisting[language=rexx,label=varvaluezero.rexx,caption=varvaluezero.rexx]{varvaluezero.rexx}

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then Var\_Value operates on the variable pool referenced by
\texttt{\#Upper} in this pool and this rule is applied to that pool.
When the pool with attribute `not-exposed' for this name is determined
the indicator returned is `D' if the name has `dropped' associated,
\texttt{\textquotesingle{}N\textquotesingle{}} otherwise.

In the former case \texttt{\#Outcome} is set equal to \texttt{Name}, in
the latter case \texttt{\#Outcome} is set to the value most recently
associated with the name by \texttt{Var\_Set}.

\lstinputlisting[language=rexx,label=varvalueone.rexx,caption=varvalueone.rexx]{varvalueone.rexx}

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then
\texttt{Var\_Value} operates on the variable pool referenced by
\texttt{\#Upper} in this pool and this rule is applied to that pool.
When the pool with attribute `not-exposed' for the stem is determined
the name is considered in that pool. If the name has the `exposed'
attribute then the variable pool referenced by \#Upper in the pool is
considered and this rule applied to that pool. When the pool with
attribute `not-exposed' is determined the indicator returned is
\texttt{\textquotesingle{}D\textquotesingle{}} if the name has `dropped'
associated, \texttt{\textquotesingle{}N\textquotesingle{}} otherwise. In
the former case \texttt{\#Outcome} is set equal to \texttt{Name}, in the
latter case \texttt{\#Outcome} is set to the value most recently
associated with the name by \texttt{Var\_Set}.

@@

\lstinputlisting[language=rexx,label=vardropzero.rexx,caption=vardropzero.rexx]{vardropzero.rexx}

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. If the specified name has the `exposed'
attribute then \texttt{Var\_Drop} operates on the variable pool
referenced by \texttt{\#Upper} in this pool and this rule is applied to
that pool. When the pool with attribute `not-exposed' for this name is
determined the attribute `dropped' is associated with the specified
name. Also, when the name is a stem,
\texttt{Var\_Drop(Pool,TailedName,\textquotesingle{}1\textquotesingle{})}
is executed for all possible valid tailed names which have \texttt{Name}
as a stem.

\lstinputlisting[language=rexx,label=vardropone.rexx,caption=vardropone.rexx]{vardropone.rexx}

The function operates on the variable pool with the specified reference.
The name is a tailed name. The left part of the name, up to and
including the first period, is the stem. The stem is a non-tailed name.
If the specified stem has the `exposed' attribute then Var\_Drop
operates on the variable pool referenced by \#Upper in this pool and
this rule is applied to that pool. When the pool with attribute
`not-exposed' for the stem is determined the name is considered in that
pool. If the name has the `exposed' attribute then the variable pool
referenced by \texttt{\#Upper} in the pool is considered and this rule
applied to that pool. When the pool with attribute `not-exposed' is
determined the attribute `dropped' is associated with the specified
name.

@@

\lstinputlisting[language=rexx,label=varexposezero.rexx,caption=varexposezero.rexx]{varexposezero.rexx}

The function operates on the variable pool with the specified reference.
The name is a non-tailed name. The attribute `exposed' is associated
with the specified name. Also, when the name is a stem,
\texttt{Var\_Expose(Pool,\ TailedName,\textquotesingle{}1\textquotesingle{})}
is executed for all possible valid tailed names which have Name as a
stem.

\lstinputlisting[language=rexx,label=varexpose1.rexx,caption=varexpose1.rexx]{varexpose1.rexx}

@@

\lstinputlisting[language=rexx,label=varresetpool.rexx,caption=varresetpool.rexx]{varresetpool.rexx}

The function operates on the variable pool with the specified reference.
It establishes the effect of subsequent \texttt{API\_Next} and
\texttt{API\_NextVariable} functions (see sections nnn and nnn). A
\texttt{Var\_Reset} is implied by any \texttt{API\_} operation other
than \texttt{API\_Next} and \texttt{API\_NextVariable}.

@@

The value of a symbol which is a \emph{NUMBER} or a \emph{CONST\_SYMBOL}
which is not a reserved symbol is the content of the appropriate token.

The value of a \emph{VAR\_SYMBOL} which is ``taken as a constant'' is
the \emph{VAR\_SYMBOL} itself, otherwise the \_VAR\_SYMBOL: identifies a
variable and its value may vary during execution.

Accessing the value of a symbol which is not ``taken as a constant''
shall result in trace output, see nnn:

\lstinputlisting[language=rexx,label=tracinglevel.rexx,caption=tracinglevel.rexx]{tracinglevel.rexx}

where \texttt{Tag} is
\texttt{\textquotesingle{}\textgreater{}L\textgreater{}\textquotesingle{}}
unless the symbol is a \emph{VAR\_SYMBOL} which, when used as an
argument to \texttt{Var\_Value}, does not yield an indicator
\texttt{\textquotesingle{}D\textquotesingle{}}. In that case, the Tag is
\texttt{\textquotesingle{}\textgreater{}V\textgreater{}\textquotesingle{}}.

@@

If \emph{VAR\_SYMBOL} does not contain a period, or contains only one
period as its last character, the value of the variable is the value
associated with \emph{VAR\_SYMBOL} in the variable pool, that is
\texttt{\#Outcome} after
\texttt{Var\_Value(Pool,\ VAR\_SYMBOL,\ \textquotesingle{}0\textquotesingle{})}.

If the indicator is \texttt{\textquotesingle{}D\textquotesingle{}},
indicating the variable has the `dropped' attribute, the
\texttt{NOVALUE} condition is raised; see nnn and nnn for exceptions to
this.

\lstinputlisting[language=rexx,label=varpoolsymbol.rexx,caption=varpoolsymbol.rexx]{varpoolsymbol.rexx}

If \emph{VAR\_SYMBOL} contains a period which is not its last character,
the value of the variable is the value associated with the derived name.

\subsection{Derived names}\label{derived-names}

A derived name is derived from a \emph{VAR\_SYMBOL} as follows:

\lstinputlisting[language=rexx,label=derivednames.rexx,caption=derivednames.rexx]{derivednames.rexx}

The derived name is the concatenation of:

\begin{itemize}
\tightlist
\item
  the \emph{Stem}, without further evaluation;
\item
  the \emph{Tail}, with the \emph{PlainSymbols} replaced by the values
  of the symbols. The value of a \emph{PlainSymbol} which does not start
  with a digit is \texttt{\#Outcome} after
\end{itemize}

\lstinputlisting[language=rexx,label=varvalueplainsymbol.rexx,caption=varvalueplainsymbol.rexx]{varvalueplainsymbol.rexx}

These values are obtained without raising the \texttt{NOVALUE}
condition.

@@

If the comparison has a comparison\_operator that is a normal\_compare
then the variable \#Test is set as follows:

\lstinputlisting[language=rexx,label=evaluation-comparison.rexx,caption=evaluation-comparison.rexx]{evaluation-comparison.rexx}

The value of \#Test, in conjunction with the operator in the comparison,
determines the value of the comparison. The value of the comparison is
`1' if - \#Test is `E' and the operator is one of `=``\,``,'==``\,``,
`\textgreater=', \textless='', `\textgreater{}', `\textless{}',
`p\textgreater=', `\textless\textless=', \textgreater\textgreater', or
\textless\textless)

\begin{itemize}
\tightlist
\item
  \#Test is `G' and the operator is one of `\textgreater{}',
  `\textgreater=``, `\textless{}', `=', `\textless\textgreater{}',
  `\textgreater\textless{}', Nes'','\textgreater\textgreater!
  `p\textgreater{}', or \textless\textless``)
\item
  \#Test is `L' and the operator is one of `\textless{}', \textless=``,
  \textgreater{}`, =', `\textless\textgreater{}',
  `\textgreater\textless{}', ==`,'\textless\textless',
  *\textless\textless=`, or \textgreater\textgreater{}'. In all other
  cases the value of the comparison is `0'.
\end{itemize}

@@

The code of \texttt{ArithOpp} itself is assumed to operate under a
sufficiently high setting of numeric digits to avoid exponential
notation.

\lstinputlisting[language=rexx,label=evaluation-arithmetic.rexx,caption=evaluation-arithmetic.rexx]{evaluation-arithmetic.rexx}

@@

Invocation occurs when a \emph{function} or a \emph{message\_term} or a
\emph{call} is evaluated. Invocation of a function may result in a
value, in which case:

\lstinputlisting[language=rexx,label=tracinglevelF.rexx,caption=tracinglevelF.rexx]{tracinglevelF.rexx}

Invocation of a \emph{message\_term} may result in a value, in which
case:

\lstinputlisting[language=rexx,label=tracinglevelM.rexx,caption=tracinglevelM.rexx]{tracinglevelM.rexx}

@@

The argument positions are the positions in the \emph{expression\_list}
where syntactically an \emph{expression} occurs or could have occurred.
Let \texttt{ArgNumber} be the number of an argument position, counting
from \texttt{1} at the left; the range of \texttt{ArgNumber} is all
whole numbers greater than zero.

For each value of \texttt{ArgNumber},
\texttt{\#ArgExists.\#NewLevel.ArgNumber} is set
\texttt{\textquotesingle{}1\textquotesingle{}} if there is an expression
present, \texttt{\textquotesingle{}O\textquotesingle{}} if not.

From the left, if \texttt{\#ArgExists.\#NewLevel.ArgNumber} is
\texttt{\textquotesingle{}1\textquotesingle{}} then
\texttt{\#Arg.\#NewLevel.ArgNumber} is set to the value of the
corresponding expression. If \texttt{\#ArgExists.\#NewLevel.ArgNumber}
is \texttt{\textquotesingle{}0\textquotesingle{}} then
\texttt{\#Arg.\#NewLevel.ArgNumber} is set to the null string.

\texttt{\#ArgExists.\#NewLevel.0} is set to the largest
\texttt{ArgNumber} for which \texttt{\#ArgExists.\#NewLevel.ArgNumber}
is \texttt{\textquotesingle{}1\textquotesingle{}}, or to zero if there
is no such value of \texttt{ArgNumber}.

@@

The value of a \emph{LABEL}, or of the \emph{taken\_constant} in the
function or \emph{call\_instruction}, is taken as a constant, see nnn.
If the \emph{taken\_constant} is not a \emph{string\_literal} it is a
reference to the first \emph{LABEL} in the program which has the same
value. The comparison is made with the `==' operator.

If there is such a matching label and the label is trace-only (see nnn)
then a condition is raised:

\lstinputlisting[language=rexx,label=callraisesyntax.rexx,caption=callraisesyntax.rexx]{callraisesyntax.rexx}

@@

If there is no such matching label, or if the \emph{taken\_constant} is
a \emph{string\_literal}, further comparisons are made.

If the value of the \emph{taken\_constant} matches the name of some
built-in function then that built-in function is invoked. The names of
the built-in functions are defined in section nnn and are in uppercase.

If the value does not match any built-in function name,
\texttt{Config\_ExternalRoutine} is used to invoke an external routine.

Whenever a matching label is found, the variables \texttt{SIGL} and
\texttt{.SIGL} are assigned the value of the line number of the clause
which caused the search for the label. In the case of an invocation
resulting from a condition occurring that shall be the clause in which
the condition occurred.

\lstinputlisting[language=rexx,label=varset.rexx,caption=varset.rexx]{varset.rexx}

The name used in the invocation is held in \texttt{\#Name.\#Level} for
possible use in an error message from the \texttt{RETURN} clause, see
nnn

@@

An internal routine completes when \texttt{\#Level} returns to the value
it had when the routine was invoked. The value of the internal function
is the value of the \emph{expression} on the \emph{return} which
completed the routine. The value of an external function is determined
by \texttt{Config\_ExternalRoutine}.

@@

An internal method completes when \texttt{\#Level} returns to the value
it had when the routine was invoked. The value of the internal method is
the value of the \emph{expression} on the \emph{return} which completed
the method. The value of an external method is determined by
\texttt{Config\_ExternalMethod}.

@@

See nnn for the syntax of a \emph{message\_term}. The value of the
\emph{term} within a \emph{message\_term} is called the receiver.

The receiver and any arguments of the term are evaluated, in left to
\lstinputlisting[language=rexx,label=evaluatemessageterm.rexx,caption=evaluatemessageterm.rexx]{evaluatemessageterm.rexx}

If the message term contains
\texttt{\textquotesingle{}\textasciitilde{}\textasciitilde{}\textquotesingle{}}
the value of the message term is the receiver.

\emph{Any effect on .Result?}

Otherwise the value of a \emph{message\_term} is the value of the method
it invokes. The method invoked is determined by the receiver and the
\emph{taken\_constant} and \emph{symbol}.

\lstinputlisting[language=rexx,label=evaluatemessagetermconstant.rexx,caption=evaluatemessagetermconstant.rexx]{evaluatemessagetermconstant.rexx}

If there is a \emph{symbol}, it is subject to a constraints.

\lstinputlisting[language=rexx,label=evaluatemessagtermsymbol.rexx,caption=evaluatemessagtermsymbol.rexx]{evaluatemessagtermsymbol.rexx}

The search will progress from the object to its class and superclasses.

\lstinputlisting[language=rexx,label=messagemethodsearch.rexx-\textgreater,caption=messagemethodsearch.rexx-\textgreater]{messagemethodsearch.rexx-\textgreater}

@@

The values of the arguments to the use of
\texttt{Config\_ExternalRoutine}, in order, are:

The argument \texttt{How} is
\texttt{\textquotesingle{}SUBROUTINE\textquotesingle{}} if the
invocation is from a \emph{call},
\texttt{\textquotesingle{}FUNCTION\textquotesingle{}} if the invocation
is from a \emph{function}.

The argument \texttt{NameType} is
\texttt{\textquotesingle{}1\textquotesingle{}} if the
\emph{taken\_constant} is a \emph{string\_literal},
\texttt{\textquotesingle{}0\textquotesingle{}} otherwise.

The argument \texttt{Name} is the value of the \emph{taken\_constant}.

The argument \texttt{Environment} is the value of this argument on the
\texttt{API\_Start} which started this execution.

The argument \texttt{Arguments} is the \texttt{\#Arg.} and
\texttt{\#ArgExists.} data.

The argument \texttt{Streams} is the value of this argument on the
\texttt{API\_Start} which started this execution.

The argument \texttt{Traps} is the value of this argument on the
\texttt{API\_Start} which started this execution.

\texttt{Var\_Reset} is invoked and \texttt{\#API\_Enabled} set to
\texttt{\textquotesingle{}1\textquotesingle{}} before use of
\texttt{Config\_ExternalRoutine}. \texttt{\#API\_Enabled} is set to
\texttt{\textquotesingle{}O\textquotesingle{}} after.

The response from \texttt{Config\_ExternalRoutine} is processed. If no
conditions are (implicitly) raised, \texttt{\#Outcome} is the value of
the function.

### 2024-04-12 ARB/standards/work-in-progress/document/tex/standard/foreword.tex
work in standard book production process including readme.md
upd
add filenames

\hypertarget{foreword}{%
\chapter{Foreword}\label{foreword}}

\hypertarget{purpose}{%
\section{Purpose}\label{purpose}}

This standard provides an unambiguous definition of the programming
language Rexx. Its purpose is to facilitate portability of Rexx programs
for use on a wide variety of computer systems. History The computer
programming language Rexx was designed by Mike Cowlishaw to satisfy the
following principal aims:

@@

The popularity of ``Object Oriented'' programming, and the need for Rexx
to work with objects created in various ways, led to Rexx extensions and
to a second X3J18 project which produced this standard. (Ed - hopefully)

\hypertarget{committee-lists}{%
\section{Committee lists}\label{committee-lists}}

(Here)

This standard was prepared by the Technical Development Committee for
Rexx, X3J18. There are annexes in this standard; they are informative
and are not considered part of this standard.

Suggestions for improvement of this standard will be welcome. They
should be sent to the

Information Technology Industry Council, 1250 Eye Street, NW, Washington
DC 20005-3922.

@@

To be inserted The people who contributed to Technical Committee J18 on
Rexx, which developed this standard, include:

>>>>

%preprocessed texin
\chapter{Foreword}\label{foreword}

\section{Purpose}\label{purpose}

This standard provides an unambiguous definition of the programming
language Rexx. Its purpose is to facilitate portability of Rexx programs
for use on a wide variety of computer systems.

\section{History}\label{history}

The computer programming language Rexx was designed by Mike Cowlishaw to
satisfy the following principal aims:

@@

The popularity of ``Object Oriented'' programming, and the need for Rexx
to work with objects created in various ways, led to Rexx extensions and
to a second X3J18 project which produced this standard. \emph{(Ed -
hopefully)}

Committee lists \emph{(Here)}

This standard was prepared by the Technical Development Committee for
Rexx, X3J18.

There are annexes in this standard; they are informative and are not
considered part of this standard.

Suggestions for improvement of this standard will be welcome. They
should be sent to the Information Technology Industry Council, 1250 Eye
Street, NW, Washington DC 20005-3922.

@@

\emph{To be inserted}

The people who contributed to Technical Committee J18 on Rexx, which
developed this standard, include:
